---
title: ""
output: 
  html_document: 
    df_print: paged
    code_folding: hide
---

Data wrangling is a set of procedures aimed at "cleaning" a data frame, removing inconsistencies, typos and mistakes, to prepare a solid ground for further analysis.

The mindset that will guide us here will be to remove rows that don't pertain to real transactions between the retailer and the customers.

Building up from our previous document, we will start by checking, for every column, if the definitions stated in the source are respected.

```{r, include = FALSE}
library(readxl)
df <- read_excel(here::here("online_retail_II.xlsx"))
library(dplyr)
df <- df %>%
  mutate(CustomerID = as.character(`Customer ID`), .keep = "unused", .after = Price)
```

<br>

# - *Invoice*

The first definition determines that, if an invoice starts with "C", like C489449, it means that it has been cancelled.

But we can also find values that start with a letter different that C.

```{r}
library(stringr)
df %>%
  filter(if_else(str_length(Invoice) == 6, FALSE,
                 if_else(str_starts(Invoice, "C"), FALSE, TRUE)))
```

Invoices that don't seem to be actual transactions with a customer, so we will remove them.

<br>

# - *StockCode*

About the stock codes, not all of them are "5-digit integral number".

```{r}
df %>%
  filter(str_length(StockCode) != 5)
```

Among those, values like 79323P and 79323W are actual transactions and must be kept.

From the remaining ones,

```{r}
df %>%
  filter(str_length(StockCode) != 5 &
           !str_detect(StockCode, "^\\d{5}[a-zA-Z]{1,2}$")) %>%
  count(StockCode, Description, sort = TRUE, name = "Number of Occurrences")
```

we will keep PADS and the values starting with DCGS, SP or gift, leaving us with 2900 rows to remove from the data frame.

```{r}
df %>%
  filter(str_length(StockCode) != 5 &
           !str_detect(StockCode, "^\\d{5}[a-zA-Z]{1,2}$") &
           !str_detect(StockCode, "PADS|DCGS|SP|gift"))
```

## - *D for Discount*

About D, that stands for Discount, we follow the common understanding that, under certain conditions, a discount is applied to an invoice to lower its total amount, and already we notice that none of them have a negative value, as one should expect, in the `Price` column.

```{r}
library(knitr)
kable(df %>%
        filter(any(StockCode == "D") &
                 Price < 0, .by = Invoice) %>%
        tally(name = "Number of Discounted Stock Codes with a Negative Price"), align = "l")
```

It could be the case that the discount is applied through the negative value in the `Quantity` column (multiplying Quantity per Price), but there are invoices (the confirmed ones) with a positive value, as we can see in the following table, showing all the invoices with a discount, 90 in total for 131 rows, a small number out of `r nrow(distinct(df, Invoice))` invoices.

```{r}
df %>% 
  group_by(Invoice) %>%
  filter(any(StockCode == "D"))
```

81 out of 90 are single line invoices, so we don't understand what they should have discounted.

```{r}
df %>% 
  group_by(Invoice) %>%
  filter(any(StockCode == "D")) %>%
  count(Invoice, sort = TRUE, name = "Number of Items")
```

Maybe it's a discount on a previous invoice, but we wouldn't know how to connect them together, plus 76 out of 81 of those single line discounted invoices has been cancelled as well.

```{r}
df %>% 
  group_by(Invoice) %>%
  filter(any(StockCode == "D") &
           n() == 1) %>%
  ungroup() %>%
  mutate("Status of Single Line Invoices" = if_else(str_starts(Invoice, "C"), "Cancelled", "Confirmed")) %>%
  count(`Status of Single Line Invoices`, name = "Number of Invoices")
```

The 5 that have not been cancelled have, like the others, a positive value in the `Price` column and that, putting aside a mistake that can be easily fixed, would beg the question on whether D (Discount) is an article that can be bought and then redeemed later, like a voucher, but we already have the gift_xxxx_xx stock code for that, unless that is specific to "Dotcomgiftshop".

```{r}
df %>% 
  group_by(Invoice) %>%
  filter(any(StockCode == "D") &
           !str_starts(Invoice, "C"))
```

About the 9 invoices with more than one line, shown here;

```{r}
df %>% 
  group_by(Invoice) %>%
  filter(any(StockCode == "D") &
           n() > 1)
```

3 of them have more lines but it is always D the stock code.

```{r}
df %>% 
  group_by(Invoice) %>%
  filter(all(StockCode == "D") &
           n() > 1)
```

The 6 remaining can be investigated further, at a first glance we can see that they all have been cancelled.

```{r}
df %>% 
  group_by(Invoice) %>%
  filter(any(StockCode == "D") &
           n() > 1) %>%
  anti_join(df %>%
              group_by(Invoice) %>%
              filter(all(StockCode == "D") &
                       n() > 1), by = "Invoice")
```

Furthermore the value of discounts in percentage is very high for some of them, it seems more like a refund plus eventual fines.

```{r}
df %>% 
  group_by(Invoice) %>%
  filter(any(StockCode == "D") &
           n() > 1) %>%
  anti_join(df %>%
              group_by(Invoice) %>%
              filter(all(StockCode == "D") &
                       n() > 1), by = "Invoice") %>%
  mutate(Status = if_else(str_detect(StockCode, "D"), "Discount", "Purchases")) %>%
  group_by(Status, .add = TRUE) %>%
  summarise(`Value in £` = sum(abs(Quantity * Price)), .groups = "drop_last") %>%
  mutate(`Discount Percentage` = formattable::percent(if_else(Status == "Discount",
                                                              `Value in £`[Status == "Discount"] / `Value in £`[Status == "Purchases"],
                                                              NA))) %>%
  arrange(desc(Status), .by_group = TRUE)
```

Considering everything, how few valuable information we could extract from the rows that have with D as stock code and that there are only 131 of them (out of `r nrow(df)`), we decide to remove them.

<br>

# - *Description*

Another discrepancy is in the different number of distinct stock codes and descriptions, that should be the same.

```{r}
df %>%
  summarise("Number of Distinct Stock Codes" = n_distinct(StockCode),
            "Number of Distinct Descriptions" = n_distinct(Description))
```

<br>

That is because some stock codes (2003 out of 4631) have several descriptions.

```{r}
df %>%
  count(StockCode, Description, name = "Number of Occurrences") %>%
  group_by(StockCode) %>%
  filter(n() > 1)
```

For 1564 out of 2003 an additional description is a missing value (NA).

```{r}
df %>%
  count(StockCode, Description, name = "Number of Occurrences") %>% 
  group_by(StockCode) %>%
  filter(n() > 1) %>%
  semi_join(df %>%
              count(StockCode, Description) %>%
              group_by(StockCode) %>%
              filter(any(is.na(Description))), by = "StockCode")
```

While the others (439) present typos, updated descriptions or notes.

```{r}
df %>%
  count(StockCode, Description, name = "Number of Occurrences") %>% 
  group_by(StockCode) %>%
  filter(n() > 1) %>%
  anti_join(df %>%
              count(StockCode, Description) %>%
              group_by(StockCode) %>%
              filter(any(is.na(Description))), by = "StockCode")
```

The latter case is identifiable as the notes are usually written in lower case.

```{r}
df %>%
  count(StockCode, Description, name = "Number of Occurrences") %>% 
  group_by(StockCode) %>%
  filter(n() > 1) %>%
  anti_join(df %>%
              count(StockCode, Description) %>%
              group_by(StockCode) %>%
              filter(any(is.na(Description))), by = "StockCode") %>%
  filter(str_detect(Description, "[:lower:]"))
```

<br>

We can also encounter the same description pertaining to different stock codes; some they just differ in style (84558A and 84558a for example) while others are two different ones (35815B and 35817B, both with ACRYLIC JEWEL SNOWFLAKE,BLUE as a description).

```{r}
df %>%
  count(Description, StockCode, name = "Number of Occurrences") %>%
  group_by(Description) %>%
  filter(n() > 1)
```

For the second instance, we discovered as well stock codes that change description.

```{r}
df %>%
  filter(StockCode %in% c("22829", "22832")) %>%
  count(StockCode, Description, name = "Number of Occurrences") #VEDERE SE RIESCO A TIRARLI FUORI SENZA IL FILTRO, PROB CON UN group_by() PRIMA
#DIREI CHE NON CE LA FACCIO, DOVREI FILTRARE TROPPA ROBA
# df %>%
#   count(StockCode, Description) %>%
#   group_by(StockCode) %>%
#   filter(n() > 1 &
#            !str_detect(Description, "[:lower:]"))
# df %>%
#   count(StockCode, Description) %>%
#   semi_join(df %>%
#               group_by(StockCode) %>%
#               filter(Description != lag(Description) &
#                        !str_detect(Description, "[:lower:]") &
#                        !str_detect(Description, "SPOT"))) %>%
#   group_by(StockCode) %>%
#   filter(n() > 1)
```

So we might say that `Description` is not a column we can rely on too much.

<br>

# - *Quantity*

Moving on to `Quantity`, let's investigate the rows with a negative value in it, which amount to 12326 of them,

```{r}
df %>%
  filter(Quantity < 0)
```

belonging to 6712 invoices, out of which 4591 are cancelled, and for them it makes sense that the quantity is negative, as a way to readjust the inventory levels of the stock code.

```{r}
df %>%
  filter(any(Quantity < 0) &
           row_number() == 1, .by = Invoice) %>%
  count(Status = if_else(str_starts(Invoice, "C"), "Cancelled", "Confirmed"), name = "Number of Invoices")
```

But the rest are not cancelled invoices and they seem to have some common concurrences,

```{r}
df %>%
  filter(Quantity < 0 &
           !str_detect(Invoice, "C"))
```

besides being all single line invoices,

```{r}
# kable(df %>%
#         semi_join(df %>%
#                     filter(Quantity < 0) %>%
#                     count(Invoice) %>%
#                     filter(!str_detect(Invoice, "C")), by = "Invoice") %>%
#         filter(n() > 1, .by = "Invoice") %>% #QUESTO E' ESEMPIO DI .by CHE OFFUSCA UN PO'
#         tally(name = "Number of Confirmed Negative Quantity Invoices with more than One Line"), align = "l")
kable(df %>%
        filter(Quantity < 0 &
                 !str_detect(Invoice, "C")) %>%
        filter(n() > 1, .by = "Invoice") %>%
        tally(name = "Number of Confirmed Negative Quantity Invoices with more than One Line"), align = "l")
```

like NAs or notes in the `Description` column,

```{r}
df %>%
  filter(Quantity < 0 &
           !str_detect(Invoice, "C")) %>%
  count(Description, sort = TRUE, name = "Number of Occurrences")
```

and with all the same value in the `Price` (0), `CustomerID` (NA) and `Country` (United Kingdom) columns. 

```{r}
df %>%
  filter(Quantity < 0 &
           !str_detect(Invoice, "C")) %>%
  distinct(Price, CustomerID, Country)
```

These invoices run through all the data frame so the issue is not time specific.

```{r}
df %>%
  filter(Quantity < 0 &
           !str_detect(Invoice, "C")) %>%
  slice(1, n())
```

We can assume that these are not actual transactions but inventory adjustments, so we will remove them.

We've seen that, when an invoice has a C, its quantity is negative but there is one case where that is not true.

```{r}
df %>%
  filter(str_detect(Invoice, "C") &
           Quantity >= 0) 
```

There are no quantity values equal to 0.

```{r}
kable(df %>%
        filter(Quantity == 0) %>%
        tally(name = "Number of Rows with 0 Quantity"), align = "l")
```

<br>

# - *InvoiceDate*

About `InvoiceDate`, let's look for gaps.
QUI CI STAREBBE BENE UN GRAFICO

```{r}
date_range <- seq(min(as.Date(df$InvoiceDate)), max(as.Date(df$InvoiceDate)), by = 1)
date_range[!date_range %in% as.Date(df$InvoiceDate)]
```

There is a big gap around Xmas 2009 and a smaller one at the beginning of April 2010 (Easter fell on the 4th of April that year) but, apart from those national holidays there are many more missing dates in our data frame. Could they pertain to a specific day of the week?

```{r}
tibble("day missing" = date_range[!date_range %in% as.Date(df$InvoiceDate)], 
       "day of the week" = weekdays(date_range[!date_range %in% as.Date(df$InvoiceDate)]))

table(factor(weekdays(date_range[!date_range %in% as.Date(df$InvoiceDate)]),
      levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"), ordered = TRUE))
```

I was expecting Sundays to be honest given that the clientele is mostly UK based. There might be other factors at play here given the nature of the business.

<br>

# - *Price*

Let's look for abnormal prices now and for the negative ones we see that they pertain to the invoices with an A.

```{r}
df %>%
  filter(Price < 0)
```

And for the ones with a price equal to 0 (3687 rows), 

```{r}
df %>%
  filter(Price == 0)
```

they seem to be a superset of the ones with a negative quantity and a not cancelled invoice (2121 rows).

```{r}
df %>%
  semi_join(df %>%
              filter(Quantity < 0) %>%
              count(Invoice) %>%
              filter(!str_detect(Invoice, "C")), by = "Invoice")
```

Within the 1566 remaining rows of this set, that have a positive Quantity,

```{r}
df %>%
  filter(Price == 0 &
           Quantity > 0)
```

we have missing values (NAs) in most of the `Description`

```{r}
df %>%
  filter(Price == 0 &
           Quantity > 0) %>%
  count(Description, sort = TRUE, name = "Number of Occurrences")
```

and `CustomerID` columns.

```{r}
df %>%
  filter(Price == 0 &
           Quantity > 0) %>%
  count(CustomerID, sort = TRUE, name = "Number of Occurrences")
```

The time frame is the same as all the data frame.

```{r}
df %>%
  filter(Price == 0 &
           Quantity > 0) %>%
  slice(1, n())
```

And they mostly pertain to United Kingdom.

```{r}
df %>%
  filter(Price == 0 &
           Quantity > 0) %>%
  count(Country, sort = TRUE)
```

<br>

# - *CustomerID*

About `CustomerID`, we have 5 customers, out of `r length(unique(df$CustomerID))`, that changed country but that is not a concern, more something to keep in mind.

```{r}
df %>%
  count(CustomerID, Country, name = "Number of Occurrences") %>%
  group_by(CustomerID) %>%
  filter(n() > 1 &
           !is.na(CustomerID))
```

<br>

# - *Country*

About `Country`, we noticed in the previous document the value "Unspecified" for 15 invoices and 5 customers, that is better to change to NA. PENSARE A COME ORDINARLO AFFINCHE' SIA PIU' CHIARO

```{r}
df %>%
  distinct(Invoice, CustomerID, Country) %>%
  filter(Country == "Unspecified") %>%
  arrange(CustomerID, Invoice)
```

<br>

# - *main takeaways*

- Invoices starting with a letter different than "C"  
- Stock codes not pertaining to actual transactions  
- A D (Discount) stock code that doesn't provide valuable information  
- A `Description` column with several values for the same stock code, missing values, typos, updated descriptions or notes, the same description assigned to different stock codes, stock codes that changed description   
- 2121 non cancelled invoices with a negative value in the `Quantity` column, all with the same value in the `Price` (0), `CustomerID` (NA) and `Country` (United Kingdom) columns. 1 cancelled invoice with a positive value in the `Quantity` column  
- Unexpected gaps in the time frame  
- 3 rows with a negative price (the same as the ones with invoices starting with a letter different than "C"). 3687 rows with price equal to 0, a set composed of rows with a negative (2121 rows) and with a positive (1566) value in the `Quantity` column 
- 5 customers that changed country  
- 15 invoices and 5 customers with an "Unspecified" value in the `Country` column  

<br>

# - *actions to be performed*

<br>

- Invoices that start with "A", not actual transactions with a customer, to be removed with

```{r class.source = "fold-show", eval = FALSE}
df %>%
  filter(!str_detect(Invoice, "A"))
```

<br>

- Stock codes not pertaining to actual transactions, to be removed with 

```{r class.source = "fold-show", eval = FALSE}
df %>%
  filter(str_length(StockCode) == 5 |
           str_detect(StockCode, "^\\d{5}[a-zA-Z]{1,2}$") |
           str_detect(StockCode, "PADS|DCGS|SP|gift"))
```

This will take care also of the invoices starting with an "A", as their stock code is not among the one preserved.

<br>

- Non cancelled invoices with a negative quantity, to be removed with

```{r class.source = "fold-show", eval = FALSE}
# df %>%
#   filter(!(Quantity < 0 & #NON NECESSARIO
#            Price == 0))
df %>%
  filter(!(Quantity < 0 &
           !str_detect(Invoice, "C")))
```

<br>

- One cancelled invoice with a positive quantity, to be removed with

```{r class.source = "fold-show", eval = FALSE}
df %>%
  filter(!(str_detect(Invoice, "C") &
             Quantity > 0))
```

But that is not necessary, as the relative stock code is not among the one preserved.

<br>

- Additional rows with a price equal to 0 and most descriptions and customerIDs empty, to be removed with

```{r class.source = "fold-show", eval = FALSE}
df %>%
  filter(!Price == 0) #QUESTO CI VUOLE
```

That will take care as well of the non cancelled invoices with a negative quantity, as those all have a price equal to 0. CI METTO UN kable() QUI?

```{r}
df %>%
  filter(Quantity < 0 &
           !str_detect(Invoice, "C")) %>%
  count("Values of Price" = Price, name = "Number of Occurrences")
```

<br>

So we can remove all non transactions rows with this code, to which we added the manipulation on the `Country` column as well (changing the "Unspecified" values to NAs).

```{r class.source = "fold-show"}
df_cleaned <- df %>%
  filter(str_length(StockCode) == 5 |
           str_detect(StockCode, "^\\d{5}[a-zA-Z]{1,2}$") |
           str_detect(StockCode, "PADS|DCGS|SP|gift")) %>%
  filter(Price != 0) %>% #VEDERE SE QUESTA POSSO UNIRLA, RIGUARDARE ANCHE LA PARTE SU my dplyr handbook PER VEDERE SE E' CHIARA ABBASTANZA
  mutate(Country = na_if(Country, "Unspecified"), .keep = "unused", .after = Price)
# df <- df %>%
#   filter(str_detect(StockCode, "[0-9]") &
#            str_length(StockCode) == 5 |
#            str_detect(StockCode, "^\\d{5}[a-zA-Z]{1,2}$") | 
#            str_detect(StockCode, "PADS|DCGS|SP|gift")) %>%
#   filter(!Price == 0) %>%
#     mutate(Country = na_if(Country, "Unspecified"), .keep = "unused", .after = Price)
```

<br>

# - *resulting modifications*

After these manipulations, the data frame has new characteristics: 

- `r nrow(df_cleaned)` rows compared to `r nrow(df)` (a difference of `r nrow(df) - nrow(df_cleaned)` rows)

- new numbers of distinct values for the character columns

```{r}
bind_rows(`Original Data Frame` = df %>%
            summarise(across(where(is.character), n_distinct)),
          `Cleaned Data Frame` = df_cleaned %>%
            summarise(across(where(is.character), n_distinct)), .id = "")
```

<br>

- a different percentage of cancelled invoices, that increased from the previous 15.94%

```{r}
df_cleaned %>%
  mutate(Status = if_else(str_starts(Invoice, "C"), "Cancelled", "Confirmed")) %>%
  group_by(Status) %>%
  summarize(`Distinct Invoices Cleaned Data Frame` = n_distinct(Invoice)) %>%
  mutate(`Cleaned Percentage` = formattable::percent(`Distinct Invoices Cleaned Data Frame` / sum(`Distinct Invoices Cleaned Data Frame`))) %>%
  arrange(desc(`Distinct Invoices Cleaned Data Frame`)) %>%
  bind_cols(df %>%
              mutate(Status = if_else(str_starts(Invoice, "C"), "Cancelled", "Confirmed")) %>%
              group_by(Status) %>%
              summarize(`Distinct Invoices Original Data Frame` = n_distinct(Invoice)) %>%
              mutate(`Original Percentage` = formattable::percent(`Distinct Invoices Original Data Frame` / sum(`Distinct Invoices Original Data Frame`)), 
                     Status = NULL) %>%
              arrange(desc(`Distinct Invoices Original Data Frame`)))
```

<br>

- and new distributions for the numeric columns

```{r}
tibble("Statistic" = c("Min." , "1st Qu.", "Median", "Mean", "3rd Qu.", "Max."))  %>%
  bind_cols(df %>%
              reframe(across(where(is.numeric), ~ summary(.x))) %>%
              rename(`Original Quantity` = Quantity, `Original Price` = Price),
            df_cleaned %>%
              reframe(across(where(is.numeric), ~ summary(.x))) %>%
              rename(`Cleaned Quantity` = Quantity, `Cleaned Price` = Price)) %>%
  relocate(Statistic, ends_with("y"), everything())
```

<br>

We also communicate that the gap between distinct stock codes and distinct descriptions widened.

```{r}
bind_rows(`Original Data Frame` = df %>%
            summarise("Number of Distinct Stock Codes" = n_distinct(StockCode),
                      "Number of Distinct Descriptions" = n_distinct(Description)),
          `Cleaned Data Frame` = df_cleaned %>%
            summarise("Number of Distinct Stock Codes" = n_distinct(StockCode),
                      "Number of Distinct Descriptions" = n_distinct(Description)), .id = "")
```

Even if we removed all missing values 

```{r}
kable(df_cleaned %>%
        filter(is.na(Description)) %>%
        tally(name = "Number of NAs in the Description Column"), align = "l")
```

and notes written in lower case. QUI CI DEVO LAVORARE ANCORA, CHE CI SONO VALIDE DESCRIPTION CON FORSE UN CARATTERE MINUSCOLO
https://stackoverflow.com/questions/62999156/how-to-filter-only-vectors-that-contain-all-uppercase-letters-in-all-the-strings

```{r}
# kable(df_cleaned %>%
#         filter(str_detect(Description, "^[^[:upper:]]+$")) %>%
#         tally(name = "Number of Notes in the Description Column"), align = "l") #negate lowercase
# kable(df_cleaned %>%
#         filter(str_detect(Description, "^[[:lower:][:space:]]+$")) %>%
#         tally(name = "Number of Notes in the Description Column"), align = "l")
# df_cleaned %>%
#         filter(str_detect(Description, "[:lower:]"))
df_cleaned %>%
  count(StockCode, Description, name = "Number of Occurrences") %>% 
  group_by(StockCode) %>%
  filter(n() > 1) %>%
  anti_join(df %>%
              count(StockCode, Description) %>%
              group_by(StockCode) %>%
              filter(any(is.na(Description))), by = "StockCode") %>%
  filter(str_detect(Description, "[:lower:]"))
```


VEDIAMO SE SULLE DATE CAMBIA QUALCOSA DOPO CHE METTO IL GRAFICO
<br>

Furthermore we removed a customer that changed country (12745), 

```{r}
df_cleaned %>%
  count(CustomerID, Country, name = "Number of Occurrences") %>%
  group_by(CustomerID) %>%
  filter(n() > 1 &
           !is.na(CustomerID))
```

that is now solely located in the United Kingdom.

```{r}
df_cleaned %>%
  filter(CustomerID == "12745") %>%
  count(CustomerID, Country, name = "Number of Occurrences")
```

<br>

And of course the distribution of the `Country` column changed.
QUESTO LO DEVO MODIFICARE DATO CHE VOGLIO MOSTRARE ANCHE IL NUOVO VALORE NA TRA I PAESI

```{r}
df %>%
  count(Country, wt = n_distinct(CustomerID), sort = TRUE, name = "Number of Customers Original Data Frame") %>%
  left_join(df_cleaned %>%
              count(Country, wt = n_distinct(CustomerID), name = "Number of Customers Cleaned Data Frame"), by = "Country")
```