---
title: ""
output: 
  html_document: 
    df_print: paged
    code_folding: hide
---

```{r, include = FALSE}
library(readxl)
df <- read_excel(here::here("online_retail_II.xlsx"))
library(dplyr)
df <- df %>%
  mutate(CustomerID = as.character(`Customer ID`), .keep = "unused", .after = Price)
library(stringr)
df <- df %>%
  filter(str_length(StockCode) == 5 |
           str_detect(StockCode, "^\\d{5}[a-zA-Z]{1,2}$") |
           str_detect(StockCode, "PADS|DCGS|SP|gift")) %>%
  filter(Price != 0) %>% #VEDERE SE QUESTA POSSO UNIRLA, RIGUARDARE ANCHE LA PARTE SU my dplyr handbook PER VEDERE SE E' CHIARA ABBASTANZA
  mutate(Country = na_if(Country, "Unspecified"), .keep = "unused", .after = Price)
```

We will now proceed to study outliers, that is those numerical values that are either much bigger or much smaller than the others.
They are very interesting to investigate as they can provide insights about particular cases that don't commonly happen and furthermore they are a way to spot typos in data entry.

<br>

We only have two numeric columns in our data frame, `Quantity` and `Price`, and this is their distribution of values,

```{r}
tibble("Statistic" = c("Min." , "1st Qu.", "Median", "Mean", "3rd Qu.", "Max.")) %>%
  bind_cols(df %>%
              reframe(across(where(is.numeric), ~ summary(.x))))
```

<br>

One definition for an outlier is when its value it's either greater/smaller than the third/first quartile plus/minus 1.5 the interquartile range.

```{r class.source = "fold-show", eval = FALSE}
Upper Outliers > 3rd Quartile + 1.5 * (3rd Quartile - 1st Quartile)
Lower Outliers < 1rd Quartile - 1.5 * (3rd Quartile - 1st Quartile)
```

So for example for this set of values, 

```{r}
(x <- c(0, 1, 2, 4, 5, 5, 5, 5, 5, 5, 5, 6, 8, 9, 10))
```

these are the upper and lower bounds, the cutoff points beyond which a value is defined as an outlier:

```{r}
`Upper Bound` <- quantile(x, probs = 0.75) + 1.5 * (quantile(x, probs = 0.75) - quantile(x, probs = 0.25))
`Lower Bound` <- quantile(x, probs = 0.25) - 1.5 * (quantile(x, probs = 0.75) - quantile(x, probs = 0.25))
tibble(`Upper Bound`, `Lower Bound`)
```

Consequently, these are the values defined as upper 

```{r}
x[x > `Upper Bound`]
```

and lower outliers

```{r}
x[x < `Lower Bound`]
```

for that set of values.

<br>

We will use this method for defining outliers as it is the most used in statistics, but with some business knowledge we could use as well more specifics values as cutoff points, like thresholds determined by how many items can fit into a box, the number of items by which the transportation costs increase and so on and so forth.

<br>

# - *preliminary inspection*

In this document then we will concentrate on the `Quantity` column but before we calculate the outliers it's better to inspect its highest

```{r}
df %>%
  arrange(desc(Quantity))
```

and lowest values to see if we spot bizarre entries.

```{r}
df %>%
  arrange(Quantity)
```

Nothing seems strange here, it seems like there aren't any typos, as the higher values look like actual purchases and the negative ones are related to a cancelled order (Invoice starting with a C, as per the definition).

<br>

# - *upper outliers*

Let's calculate the upper bound for the outliers then, but before we will remove the rows with a negative value in the `Quantity` column, as those purchases have been cancelled so we feel that it would be wrong to consider them in the calculation.

```{r}
df_outlier_Quantity <- df %>%
  filter(Quantity >= 0)
`Quantity Upper Bound` <- unname(quantile(df_outlier_Quantity$Quantity, probs = 0.75) + 1.5 * (quantile(df_outlier_Quantity$Quantity, probs = 0.75) - quantile(df_outlier_Quantity$Quantity, probs = 0.25)))
library(knitr)
kable(tibble(`Quantity Upper Bound`), align = "l") 
```

So we obtained a upper bound value of 26 units, let's see then how many purchases are upper outliers in terms of quantity, 

```{r}
df_outlier_Quantity %>%
  mutate("Upper Outlier" = if_else(Quantity > `Quantity Upper Bound`, TRUE, FALSE)) %>%
  summarise("Total Number of Purchases" = n(),
            "Number of Upper Outliers Purchases" = sum(`Upper Outlier`),
            "Percentage of Upper Outliers Purchases" = formattable::percent(mean(`Upper Outlier`)))
```

and what items are more frequently sold in quantities higher than the upper bound.
RIFLETTERE SU QUALI INFORMAZIONI SIANO IMPORTANTI DA METTERE, PER OGNI TABELLA

```{r}
df_outlier_Quantity %>%
  mutate("Upper Outlier" = if_else(Quantity > `Quantity Upper Bound`, TRUE, FALSE)) %>%
  group_by(StockCode, Description) %>%
  summarise("Median Quantity" = round(median(Quantity), 0),
            "Median Price" = round(median(Price), 2),
            "Number of Purchases" = n(),
            "Number of Upper Outliers Purchases" = sum(`Upper Outlier`),
            "Percentage of Upper Outliers" = formattable::percent(mean(`Upper Outlier`)), .groups = "drop") %>%
  arrange(desc(`Percentage of Upper Outliers`))
```

Let's investigate now if the items always sold in higher quantities have some common characteristic and, starting with their price, we notice that their median price is much lower than the one of the other items.

```{r}
df_outlier_Quantity %>%
  mutate("Upper Outlier" = if_else(Quantity > `Quantity Upper Bound`, TRUE, FALSE)) %>%
  group_by(StockCode, Description) %>%
  mutate("Percentage of Upper Outliers" = mean(`Upper Outlier`),
         "100% Upper Outlier" = if_else(`Percentage of Upper Outliers` == 1, TRUE, FALSE)) %>%
  ungroup() %>%
  count(`100% Upper Outlier`, wt = round(median(Price), 2), name = "Median Price")
```

We can have a visual confirmation by showing the distribution of the `Price` column for both cases
E' UNA COSA DIVERSA QUI PERO', NON C'E' L'AGGREGAZIONE SU STOCKCODES COME SOPRA
POI SPEZZA UN PO' IL FLUSSO CON IL DISCORSO SEGUENTE

```{r}
library(ggplot2)
df_outlier_Quantity %>%
  mutate("Upper Outlier" = if_else(Quantity > `Quantity Upper Bound`, TRUE, FALSE)) %>%
  ggplot(aes(`Upper Outlier`, Price)) + 
  geom_boxplot()
```

It seems so, as for big bulks the most expensive item we have costs 52.78, 

```{r}
df_outlier_Quantity %>%
  filter(Quantity > `Quantity Upper Bound`) %>%
  summarize(highestPrice = max(Price))
```

while for smaller bulks we see from the graph many values more expensive than that.

This table shows how many distinct items we have for every percentage of upper outliers; for 100% we have 43, a small number, but not negligible.

```{r}
df_outlier_Quantity %>%
  mutate("Upper Outlier" = if_else(Quantity > `Quantity Upper Bound`, TRUE, FALSE)) %>%
  group_by(StockCode, Description) %>%
  summarise("Percentage of Upper Outliers" = mean(`Upper Outlier`), .groups = "drop") %>%
  count(`Percentage of Upper Outliers`, name = "Number of Distinct Items", sort = TRUE) %>%
  mutate(`Percentage of Upper Outliers` = formattable::percent(`Percentage of Upper Outliers`),
         "Percentage of Distinct Items" = formattable::percent(`Number of Distinct Items` / sum(`Number of Distinct Items`)))
```

As we can see as well from the following histogram, where we grouped `Percentage of Upper Outliers` in 100 bins, one for every integer percentage.

```{r}
df_outlier_Quantity %>%
  mutate("Upper Outlier" = if_else(Quantity > `Quantity Upper Bound`, TRUE, FALSE)) %>%
  group_by(StockCode, Description) %>%
  summarise("Percentage of Upper Outliers" = mean(`Upper Outlier`), .groups = "drop") %>%
  #count(`Percentage of Upper Outliers`, name = "Number of Items", sort = TRUE) %>% #PROVARE A FARLO SENZA count() E CON geom_bar()
  ggplot(aes(`Percentage of Upper Outliers`)) +
  geom_histogram(bins = 100) # QUI POI DEVO CAMBIARE LA SCALA E METTERE LE PERCENTUALI
```

We can then analogously proceed to determine if there are some days where the purchases of high quantities is more common,

```{r}
df_outlier_Quantity %>%
  mutate("Upper Outlier" = if_else(Quantity > `Quantity Upper Bound`, TRUE, FALSE)) %>%
  group_by(InvoiceDay = as.Date(InvoiceDate)) %>%
  summarise("Median Quantity" = round(median(Quantity), 0),
            "Number of Purchases" = n(),
            "Number of Upper Outliers Purchases" = sum(`Upper Outlier`),
            "Percentage of Upper Outliers" = formattable::percent(mean(`Upper Outlier`))) %>%
  arrange(desc(`Percentage of Upper Outliers`))
```

and we notice higher percentages with the beginning of the New Year and somewhere in August, but those higher percentages are not extremely higher than the rest (POI CI METTO higher than the average QUANDO METTO LA LINEA A META', METTERE LA SCALA Y FINO A 100%?).

```{r}
df_outlier_Quantity %>%
  mutate("Upper Outlier" = if_else(Quantity > `Quantity Upper Bound`, TRUE, FALSE)) %>%
  group_by(InvoiceDay = as.Date(InvoiceDate)) %>%
  summarise("Percentage of High Outliers" = formattable::percent(mean(`Upper Outlier`))) %>%
  ggplot(aes(InvoiceDay, `Percentage of High Outliers`)) +
  geom_col()
 # geom_line(aes(InvoiceDay, mean(.x$`Percentage of High Outliers`))) #COME MAI QUESTO NON FUNZIONA?
```

We can move now to customers, to identify the ones that buys in high quantities more often,

```{r}
df_outlier_Quantity %>%
  mutate("Upper Outlier" = if_else(Quantity > `Quantity Upper Bound`, TRUE, FALSE)) %>%
  group_by(CustomerID) %>%
  summarise("Median Quantity" = round(median(Quantity), 0),
            "Number of Purchases" = n(),
            "Number of Upper Outliers Purchases" = sum(`Upper Outlier`),
            "Percentage of Upper Outliers" = formattable::percent(mean(`Upper Outlier`))) %>%
  arrange(desc(`Percentage of Upper Outliers`), desc(`Number of Purchases`))
```

and how many of them there are for every percentage, both as a table

```{r}
df_outlier_Quantity %>%
  mutate("Upper Outliers" = if_else(Quantity > `Quantity Upper Bound`, TRUE, FALSE)) %>%
  group_by(CustomerID) %>%
  summarise("Percentage of Upper Outliers" = formattable::percent(mean(`Upper Outliers`))) %>%
  count(`Percentage of Upper Outliers`, name = "Number of Customers", sort = TRUE)
```

and as a graph.

```{r}
df_outlier_Quantity %>%
  mutate("Upper Outliers" = if_else(Quantity > `Quantity Upper Bound`, TRUE, FALSE)) %>%
  group_by(CustomerID) %>%
  summarise("Percentage of Upper Outliers" = mean(`Upper Outliers`)) %>%
  ggplot(aes(`Percentage of Upper Outliers`)) +
  geom_histogram(bins = 100)
```

Lastly, we can investigate the countries but we don't notice any that predominantly buys only larger quantities.

```{r}
df_outlier_Quantity %>%
  mutate("Upper Outlier" = if_else(Quantity > `Quantity Upper Bound`, TRUE, FALSE)) %>%
  group_by(Country) %>%
  summarise("Number of Purchases" = n(),
            "Number of Upper Outliers Purchases" = sum(`Upper Outlier`),
            "Percentage of Upper Outliers" = formattable::percent(mean(`Upper Outlier`))) %>%
  arrange(desc(`Percentage of Upper Outliers`))
```

<br>

# - *lower outliers*

Moving to the lower outliers, when we apply the formula it returns a negative value,

```{r}
`Quantity Lower Bound` <- unname(quantile(df_outlier_Quantity$Quantity, probs = 0.25) +-1.5 * (quantile(df_outlier_Quantity$Quantity, probs = 0.75) - quantile(df_outlier_Quantity$Quantity, probs = 0.25)))
kable(tibble(`Quantity Lower Bound`), align = "l") 
```

so we will pick an extreme value of `Quantity` equal to 1 to investigate purchases of low quantity. QUESTO POTREBBE ESSERE SIMILE A INVESTIGATING ITEMS MENO POPOLARI?

Let's proceed then as we did with the upper outliers, first looking at how many of these unitary quantity purchases there are.

```{r}
df_outlier_Quantity %>%
  mutate("Unitary Quantity" = if_else(Quantity == 1, TRUE, FALSE)) %>%
  summarise("Total Number of Purchases" = n(),
            "Total Number of Unitary Quantity Purchases" = sum(`Unitary Quantity`),
            "Percentage of Unitary Quantity Purchases" = formattable::percent(mean(`Unitary Quantity`)))
```

And then what items are most commonly sold in this amount of 1.

```{r}
df_outlier_Quantity %>%
  mutate("Unitary Quantity" = if_else(Quantity == 1, TRUE, FALSE)) %>%
  group_by(StockCode, Description) %>%
  summarise("Median Quantity" = round(median(Quantity), 0),
            "Median Price" = round(median(Price), 2),
            "Number of Purchases" = n(),
            "Number of Unitary Quantity Purchases" = sum(`Unitary Quantity`),
            "Percentage of Unitary Quantity Purchases" = formattable::percent(mean(`Unitary Quantity`)), .groups = "drop") %>%
  arrange(desc(`Percentage of Unitary Quantity Purchases`))
```

As before, their median price is different from the other items, this time being higher.

```{r}
df_outlier_Quantity %>%
  mutate("Unitary Quantity" = if_else(Quantity == 1, TRUE, FALSE)) %>%
  group_by(StockCode, Description) %>%
  mutate("Percentage of Unitary Quantity Purchases" = mean(`Unitary Quantity`),
         "100% Unitary Quantity Purchase" = if_else(`Percentage of Unitary Quantity Purchases` == 1, TRUE, FALSE)) %>%
  ungroup() %>%
  count(`100% Unitary Quantity Purchase`, wt = round(median(Price), 2), name = "Median Price")
```

About the prices, the distribution of them for a single item purchase is not different from the others.
COME SOPRA, VEDERE SE TENERLO QUESTO

```{r}
df_outlier_Quantity %>%
  mutate(Quantity_of_1 = if_else(Quantity == 1, TRUE, FALSE)) %>%
  ggplot(aes(Quantity_of_1, Price)) + 
  geom_boxplot()
```

The distribution is very different here though RIFLETTERE E AGGIUNGERE QUALCOSA

```{r}
df_outlier_Quantity %>%
  mutate("Unitary Quantity" = if_else(Quantity == 1, TRUE, FALSE)) %>%
  group_by(StockCode, Description) %>%
  summarise("Percentage of Unitary Quantity Purchases" = mean(`Unitary Quantity`), .groups = "drop") %>%
  count(`Percentage of Unitary Quantity Purchases`, name = "Number of Items", sort = TRUE) %>%
  mutate(`Percentage of Unitary Quantity Purchases` = formattable::percent(`Percentage of Unitary Quantity Purchases`),
         "Percentage of Items" = formattable::percent(`Number of Items` / sum(`Number of Items`)))
df_outlier_Quantity %>%
  mutate("Unitary Quantity" = if_else(Quantity == 1, TRUE, FALSE)) %>%
  group_by(StockCode, Description) %>%
  summarise("Percentage of Unitary Quantity Purchases" = mean(`Unitary Quantity`), .groups = "drop") %>%
  ggplot(aes(`Percentage of Unitary Quantity Purchases`)) +
  geom_histogram(bins = 100) # QUI POI DEVO CAMBIARE LA SCALA E METTERE LE PERCENTUALI
```

Let's see the days now, and we notice a spike before the Christmas holidays. CAMBIARE LE SCALE QUI, ANCHE IN QUELLO PRIMA

```{r}
df_outlier_Quantity %>%
  mutate("Unitary Quantity" = if_else(Quantity == 1, TRUE, FALSE)) %>%
  group_by(InvoiceDay = as.Date(InvoiceDate)) %>%
  summarise("Median Quantity" = round(median(Quantity), 0),
            "Number of Purchases" = n(),
            "Number of Unitary Quantity Purchases" = sum(`Unitary Quantity`),
            "Percentage of Unitary Quantity Purchases" = formattable::percent(mean(`Unitary Quantity`))) %>%
  arrange(desc(`Percentage of Unitary Quantity Purchases`))
df_outlier_Quantity %>%
  mutate("Unitary Quantity" = if_else(Quantity == 1, TRUE, FALSE)) %>%
  group_by(InvoiceDay = as.Date(InvoiceDate)) %>%
  summarise("Percentage of Unitary Quantity Purchases" = formattable::percent(mean(`Unitary Quantity`)))  %>%
  ggplot(aes(InvoiceDay, `Percentage of Unitary Quantity Purchases`)) +
  geom_col()
```

We can move now to customers, to identify the ones that buys in unitary quantities more often,

```{r}
df_outlier_Quantity %>%
  mutate("Unitary Quantity" = if_else(Quantity == 1, TRUE, FALSE)) %>%
  group_by(CustomerID) %>%
  summarise("Median Quantity" = round(median(Quantity), 0),
            "Number of Purchases" = n(),
            "Number of Unitary Quantity Purchases" = sum(`Unitary Quantity`),
            "Percentage of Unitary Quantity Purchases" = formattable::percent(mean(`Unitary Quantity`))) %>%
  arrange(desc(`Percentage of Unitary Quantity Purchases`), desc(`Number of Purchases`))
```

and how many of them there are for every percentage.

```{r}
df_outlier_Quantity %>%
  mutate("Unitary Quantity" = if_else(Quantity == 1, TRUE, FALSE)) %>%
  group_by(CustomerID) %>%
  summarise("Percentage of Unitary Quantity Purchases" = mean(`Unitary Quantity`)) %>%
  count(`Percentage of Unitary Quantity Purchases`, name = "Number of Customers", sort = TRUE)
df_outlier_Quantity %>%
  mutate("Unitary Quantity" = if_else(Quantity == 1, TRUE, FALSE)) %>%
  group_by(CustomerID) %>%
  summarise("Percentage of Unitary Quantity Purchases" = mean(`Unitary Quantity`)) %>%
  ggplot(aes(`Percentage of Unitary Quantity Purchases`)) +
  geom_histogram(bins = 100)
# df %>%
#   mutate(big_bulks = if_else(Quantity > `Quantity Upper Bound`, TRUE, FALSE)) %>%
#   count(CustomerID, big_bulks, name = "n_stockcodes") %>%
#   group_by(CustomerID) %>%
#   mutate(perc_big = formattable::percent(n_stockcodes / sum(n_stockcodes))) %>%
#   ungroup() %>%
#   filter(big_bulks == TRUE) %>%
#   count(perc_big, sort = TRUE) %>%
#   mutate(perc_100 = formattable::percent(n / sum(n))) #They are actually a lot, one fifth of the total. CONTROLLARE QUESTO COMMENTO
```

Lastly, we can investigate which countries buy in unitary quantities and we don't see any that does that predominantly, bar for Nigeria.

```{r}
df_outlier_Quantity %>%
  mutate("Unitary Quantity" = if_else(Quantity == 1, TRUE, FALSE)) %>%
  group_by(Country) %>%
  summarise("Number of Purchases" = n(),
            "Number of Unitary Quantity Purchases" = sum(`Unitary Quantity`),
            "Percentage of Unitary Quantity Purchases" = formattable::percent(mean(`Unitary Quantity`))) %>%
  arrange(desc(`Percentage of Unitary Quantity Purchases`))
```

<br>

# - *actions performed*

We didn't find any extremely large or extremely small value that could be attributed to typos or any other kind of inconveniences.

<br>

# - *main takeaways*

We built an understanding about purchases that are either much greater or much smaller than the rest in respect to `Quantity` values, identifying the items more commonly belonging to those purchases, the clients responsible for those and which days they are more common to happen.
No specific instances worth of a deeper investigation have been found though.