---
title: ""
output: 
  html_document: 
    df_print: paged
    code_folding: hide
---

```{r, include = FALSE}
library(readxl)
df <- read_excel(here::here("online_retail_II.xlsx"))
library(dplyr)
df <- df %>%
  mutate(CustomerID = as.character(`Customer ID`), .keep = "unused", .after = Price)
library(stringr)
df <- df %>%
  filter(str_length(StockCode) == 5 |
           str_detect(StockCode, "^\\d{5}[a-zA-Z]{1,2}$") |
           str_detect(StockCode, "PADS|DCGS|SP|gift")) %>%
  filter(Price != 0) %>% #VEDERE SE QUESTA POSSO UNIRLA, RIGUARDARE ANCHE LA PARTE SU my dplyr handbook PER VEDERE SE E' CHIARA ABBASTANZA
  mutate(Country = na_if(Country, "Unspecified"), .keep = "unused", .after = Price)
```

NON CI SONO RIGHE DA TOGLIERE?
After having manipulated our data frame in the previous section of the site, removing rows that are not pertinent transactions and exploring missing values, duplication and limit cases, we will here start to interrogate the data, using standard data analysis techniques, with the goal to extract information that can be used for business purposes.

<br>

# - *focus of this document*

In this document we will investigate the invoices, with a focus on the phenomenon of the cancelled ones. Cancelled invoices can be very concerning as they represent lost sales so we will try to understand how widespread they are and to quantify the lost revenues.

<br>

# - *basic breakdown*

We can begin by reminding their basic figures: how many there are and how many of them have been cancelled.

```{r}
df %>%
  mutate(Status = if_else(str_starts(Invoice, "C"), "Cancelled", "Confirmed")) %>%
  group_by(Status) %>%
  summarize(`Number of Distinct Invoices` = n_distinct(Invoice)) %>%
  mutate(Percentage = formattable::percent(`Number of Distinct Invoices` / sum(`Number of Distinct Invoices`))) %>%
  arrange(desc(`Number of Distinct Invoices`))
df %>%
  summarise(`Number of Distinct Invoices` = n_distinct(Invoice), #NON E' BRUTTISSIMO QUESTO
            `Number of Cancelled Distinct Invoices` = n_distinct(Invoice[str_starts(Invoice, "C")]),
            Percentage = formattable::percent(`Number of Cancelled Distinct Invoices` / `Number of Distinct Invoices`))
```

And then we can continue by showing the number of invoices per day, distinguishing by status.

```{r}
df %>%
  mutate(Status = if_else(str_detect(Invoice, "C"), "Cancelled", "Confirmed")) %>%
  distinct(Invoice, InvoiceDay = as.Date(InvoiceDate), Status) %>%
  count(InvoiceDay, Status, name = "Number of Invoices per Day")
library(ggplot2)
df %>%
  mutate(Status = if_else(str_detect(Invoice, "C"), "Cancelled", "Confirmed")) %>%
  distinct(Invoice, InvoiceDay = as.Date(InvoiceDate), Status) %>%
  count(InvoiceDay, Status) %>%
  ggplot(aes(InvoiceDay, n, color = Status)) +
  geom_line()
```

From the graph it seems like the cancelled invoices are staying constant while the confirmed ones are on an upward trend. RISCRIVERE DOPO CHE PLOTTO I TRENDS

We complete the previous information with their distribution of invoices per day, both with a table and with a graph, to accentuate the difference in magnitude between them.
DEVO SPECIFICARE CHE IL NUMERO DI VALORI PER LE DUE DISTRIBUZIONI E' DIVERSO, MA FORSE NON E' CORRETTO? SE IN UN GIORNO NON CI SONO CANCELLED INVOICES COME LO CONSIDERIAMO? DIREI CHE E' CORRETTO NON INSERIRE RIGHE CON CANCELLED UGUALE A 0.

```{r}
# tibble("Statistic" = c("Min." , "1st Qu.", "Median", "Mean", "3rd Qu.", "Max.")) %>%
#   bind_cols(df %>%
#               mutate(Status = if_else(str_detect(Invoice, "C"), "Cancelled", "Confirmed")) %>%
#               distinct(Invoice, InvoiceDay = as.Date(InvoiceDate), Status) %>%
#               count(InvoiceDay, Status) %>%
#               tidyr::pivot_wider(names_from = "Status", values_from = "n", values_fill = 0) %>% #QUELLO VECCHIO
#               reframe(across(where(is.numeric), ~ summary(.x))))
tibble("Statistic" = c("Min." , "1st Qu.", "Median", "Mean", "3rd Qu.", "Max.")) %>%
  bind_cols(df %>%
              mutate(Status = if_else(str_detect(Invoice, "C"), "Cancelled", "Confirmed")) %>%
              distinct(Invoice, InvoiceDay = as.Date(InvoiceDate), Status) %>%
              count(InvoiceDay, Status) %>%
              filter(Status == "Confirmed") %>%
              tidyr::pivot_wider(names_from = "Status", values_from = "n") %>% #CON cur_group() MAGARI VIENE MEGLIO
              reframe(across(where(is.numeric), ~ summary(.x))),
            df %>%
              mutate(Status = if_else(str_detect(Invoice, "C"), "Cancelled", "Confirmed")) %>%
              distinct(Invoice, InvoiceDay = as.Date(InvoiceDate), Status) %>%
              count(InvoiceDay, Status) %>%
              filter(Status == "Cancelled") %>%
              tidyr::pivot_wider(names_from = "Status", values_from = "n") %>% #
              reframe(across(where(is.numeric), ~ summary(.x))))
df %>%
  mutate(Status = if_else(str_detect(Invoice, "C"), "Cancelled", "Confirmed")) %>% #FORSE NON PROPRIO CHIARO QUESTO PER CHI NON CONOSCE STATISTICA
  distinct(Invoice, InvoiceDay = as.Date(InvoiceDate), Status) %>%
  count(InvoiceDay, Status) %>%
  ggplot(aes(n, color = Status)) +
  geom_density()
```

# - *number of distinct items*

Then we want to know how many distinct items every invoice has, always distinguishing by status, where we can see that cancelled invoices are generally smaller.

```{r}
df %>%
  mutate(Status = if_else(str_detect(Invoice, "C"), "Cancelled", "Confirmed"),
         InvoiceDay = as.Date(InvoiceDate)) %>%
  #group_by(InvoiceDay = as.Date(InvoiceDate), Invoice, Status) %>%
  summarise(`Number of Distinct Items` = n_distinct(StockCode), .by = c(InvoiceDay, Invoice, Status)) #CON .by PER AVERE L'ORDINE DELLE RIGHE ORIGINALE
tibble("Statistic" = c("Min." , "1st Qu.", "Median", "Mean", "3rd Qu.", "Max.")) %>%
  bind_cols(df %>%
              mutate(Status = if_else(str_detect(Invoice, "C"), "Cancelled", "Confirmed")) %>%
              group_by(Invoice, Status) %>%
              summarise(`Number of Distinct Items` = n_distinct(StockCode), .groups = "drop") %>%
  filter(Status == "Confirmed") %>%
              tidyr::pivot_wider(names_from = "Status", values_from = "Number of Distinct Items") %>% #CON cur_group() MAGARI VIENE MEGLIO
              reframe(across(where(is.numeric), ~ summary(.x))),
  df %>%
              mutate(Status = if_else(str_detect(Invoice, "C"), "Cancelled", "Confirmed")) %>%
              group_by(Invoice, Status) %>%
              summarise(`Number of Distinct Items` = n_distinct(StockCode), .groups = "drop") %>%
  filter(Status == "Cancelled") %>%
              tidyr::pivot_wider(names_from = "Status", values_from = "Number of Distinct Items") %>%
              reframe(across(where(is.numeric), ~ summary(.x))))
  # df %>%
  #             mutate(Status = if_else(str_detect(Invoice, "C"), "Cancelled", "Confirmed")) %>%
  #             group_by(Invoice, Status) %>%
  #             summarise(`Number of Distinct Items` = n_distinct(StockCode), .groups = "drop") %>%
  #             tidyr::pivot_wider(names_from = "Status", values_from = "Number of Distinct Items", values_fill = 0) %>% #QUELLO VECCHIO
  #             reframe(across(where(is.numeric), ~ summary(.x)))
df %>%
  mutate(Status = if_else(str_detect(Invoice, "C"), "Cancelled", "Confirmed")) %>%
  group_by(Invoice, Status) %>%
  summarise(`Number of Distinct Items` = n_distinct(StockCode), .groups = "drop") %>%
  ggplot(aes(Status, `Number of Distinct Items`)) +
  geom_boxplot() +
  scale_y_log10()
```

# - *differences in Quantity*

We proceed analogously for Quantity 

```{r}
df %>%
  mutate(Status = if_else(str_detect(Invoice, "C"), "Cancelled", "Confirmed"),
         InvoiceDay = as.Date(InvoiceDate)) %>%
  #group_by(, Invoice, Status) %>%
  summarise(`Total Quantity` = sum(Quantity), .by = c(InvoiceDay, Invoice, Status))
  #arrange(desc(`Total Quantity`))
tibble("Statistic" = c("Min." , "1st Qu.", "Median", "Mean", "3rd Qu.", "Max.")) %>%
  bind_cols(df %>%
              mutate(Status = if_else(str_detect(Invoice, "C"), "Cancelled", "Confirmed")) %>%
              group_by(Invoice, Status) %>%
              summarise(`Total Quantity` = sum(Quantity), .groups = "drop") %>%
  filter(Status == "Confirmed") %>%
              tidyr::pivot_wider(names_from = "Status", values_from = "Total Quantity") %>% #CON cur_group() MAGARI VIENE MEGLIO
              reframe(across(where(is.numeric), ~ summary(.x))),
  df %>%
              mutate(Status = if_else(str_detect(Invoice, "C"), "Cancelled", "Confirmed")) %>%
              group_by(Invoice, Status) %>%
              summarise(`Total Quantity` = sum(abs(Quantity)), .groups = "drop") %>%
  filter(Status == "Cancelled") %>%
              tidyr::pivot_wider(names_from = "Status", values_from = "Total Quantity") %>%
              reframe(across(where(is.numeric), ~ summary(.x))))
  # df %>%
  #             mutate(Status = if_else(str_detect(Invoice, "C"), "Cancelled", "Confirmed")) %>%
  #             group_by(Invoice, Status) %>%
  #             summarise(`Total Quantity` = sum(Quantity), .groups = "drop") %>%
  #             tidyr::pivot_wider(names_from = "Status", values_from = "Total Quantity", values_fill = 0) %>% #QUELLO VECCHIO
  #             reframe(across(where(is.numeric), ~ summary(.x)))
df %>%
  mutate(Status = if_else(str_detect(Invoice, "C"), "Cancelled", "Confirmed")) %>%
  group_by(Invoice, Status) %>%
  summarise(`Total Quantity` = sum(abs(Quantity)), .groups = "drop") %>%
  ggplot(aes(Status, `Total Quantity`)) +
  geom_boxplot() +
  scale_y_log10()
```

And we can see that cancelled invoices are, again, generally smaller.

C524235, 518505, 524174 COINCIDENZA CHE QUESTE TRE INVOICE HANNO LO STESSO VALORE DI TOTAL QUANTITY?

```{r}
df %>%
  filter(Invoice %in% c("C524235", "518505", "524174"))
```


# - *differences in revenues*

Besides being comprising of less distinct items, in smaller quantities, cancelled invoices are of smaller monetary value as well.

```{r}
df %>%
  mutate(Status = if_else(str_detect(Invoice, "C"), "Cancelled", "Confirmed"),
         InvoiceDay = as.Date(InvoiceDate)) %>%
  #group_by(InvoiceDay = as.Date(InvoiceDate), Invoice, Status) %>%
  summarise(`Total Value` = sum(Quantity * Price), .by = c(InvoiceDay, Invoice, Status))
  #arrange(desc(`Total Value`))
tibble("Statistic" = c("Min." , "1st Qu.", "Median", "Mean", "3rd Qu.", "Max.")) %>%
  bind_cols(df %>%
              mutate(Status = if_else(str_detect(Invoice, "C"), "Cancelled", "Confirmed")) %>%
              group_by(Invoice, Status) %>%
              summarise(`Total Value` = sum(Quantity * Price), .groups = "drop") %>%
  filter(Status == "Confirmed") %>%
              tidyr::pivot_wider(names_from = "Status", values_from = "Total Value") %>% #CON cur_group() MAGARI VIENE MEGLIO
              reframe(across(where(is.numeric), ~ summary(.x))),
  df %>%
              mutate(Status = if_else(str_detect(Invoice, "C"), "Cancelled", "Confirmed")) %>%
              group_by(Invoice, Status) %>%
              summarise(`Total Value` = sum(abs(Quantity * Price)), .groups = "drop") %>%
  filter(Status == "Cancelled") %>%
              tidyr::pivot_wider(names_from = "Status", values_from = "Total Value") %>%
              reframe(across(where(is.numeric), ~ summary(.x))))
  # df %>%
  #             mutate(Status = if_else(str_detect(Invoice, "C"), "Cancelled", "Confirmed")) %>%
  #             group_by(Invoice, Status) %>%
  #             summarise(`Total Quantity` = sum(Quantity), .groups = "drop") %>%
  #             tidyr::pivot_wider(names_from = "Status", values_from = "Total Quantity", values_fill = 0) %>% #QUELLO VECCHIO
  #             reframe(across(where(is.numeric), ~ summary(.x)))
df %>%
  mutate(Status = if_else(str_detect(Invoice, "C"), "Cancelled", "Confirmed")) %>%
  group_by(Invoice, Status) %>%
  summarise(`Total Value` = sum(abs(Quantity * Price)), .groups = "drop") %>%
  ggplot(aes(Status, `Total Value`)) +
  geom_boxplot() +
  scale_y_log10()
```

The total loss in revenues can be quantified to 2.46%. RIFLETTERE SU COME CALCOLARE LA PERCENTUALE, DEVONO ESSERE I SOLDI CHE NON ENTRANO

```{r}
df %>%
  mutate(Status = if_else(str_detect(Invoice, "C"), "Cancelled", "Confirmed")) %>%
  group_by(Status) %>%
  summarise(`Total Value` = sum(abs(Quantity) * Price), .groups = "drop") 
df %>%
  summarise(`Total Value Confirmed Invoices` = sum(abs(Quantity[!str_starts(Invoice, "C")]) * Price[!str_starts(Invoice, "C")]), #QUESTO INVECE NON E' BELLISSIMO
            `Total Value Cancelled Invoices` = sum(abs(Quantity[str_starts(Invoice, "C")]) * Price[str_starts(Invoice, "C")]),
            Percentage = formattable::percent(`Total Value Cancelled Invoices` / (`Total Value Confirmed Invoices` + `Total Value Cancelled Invoices`)))
  
```



# - *country breakdown*

Let's look now at the number of invoices per country, where we see that more than 9 invoices out of 10 pertain to the United Kingdom.

```{r}
df %>%
  count(Country, sort = TRUE, name = "Number of Occurrences") %>%
  mutate(Percentage = formattable::percent(`Number of Occurrences` / sum(`Number of Occurrences`)))
df %>%
  count(Country) %>%
  ggplot(aes(reorder(Country, n), n)) +
  geom_col() +
  coord_flip()
```

More interesting to see if some countries cancel invoices more often than others, and we can spot Japan and Korea with fairly higher cancellation rates.

```{r}
df %>%
  mutate(Status = if_else(str_detect(Invoice, "C"), "Cancelled", "Confirmed")) %>%
  count(Country, Status, name = "Number of Occurrences") %>%
  group_by(Country) %>%
  mutate(Percentage = formattable::percent(`Number of Occurrences` / sum(`Number of Occurrences`)),
         Arr_Column = sum(`Number of Occurrences`)) %>%
  arrange(desc(Arr_Column)) %>% #VEDERE SE QUESTO POSSO FARLO MEGLIO, MAGARI CON cur_group()?
  select(-Arr_Column)
library(forcats)
df %>%
  mutate(Status = if_else(str_detect(Invoice, "C"), "Cancelled", "Confirmed")) %>%
  count(Country, Status) %>%
  group_by(Country) %>%
  mutate(perc = formattable::percent(n / sum(n))) %>%
  ggplot(aes(reorder(Country, n), perc, fill = fct_rev(Status))) +
  geom_col() +
  coord_flip()
```

# - *main takeaways*

- Cancelled invoices, being 16.64% of the total, are a not trascurable phenomenon.
- Their daily occurrence stayed constant during last year though while confirmed invoices are on a rising trend.
- They are generally smaller in terms of distinct items, total quantity and revenue compared to the confirmed ones.
- The loss in revenue can be quantified to 247901.4	£, the 2.46%.
- The market is mainly based in the United Kingdom, amongst other countries Japan and Korea are interesting as they have a much higher cancellation rate.